---
title: "STAT 534 - Lecture 23:"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.align= 'center')
knitr::opts_chunk$set(fig.height= 4)
knitr::opts_chunk$set(fig.width = 6)
library(tidyverse)
library(gridExtra)
library(ggplot2)
library(smacpod)
library(spatstat)
set.seed(04082019)
```

### K function

- We previously looked at the $F(d)$ and $G(d)$ functions, which corresponded to 

\vfill

- Another interesting feature of a point process is the number of points in a specified area.
\vfill
- Consider $E(Num(\boldsymbol{s},d,\boldsymbol{S}))$, the expected number of points in $\delta_d\boldsymbol{s}$, a circle of radius $d$ centered at $\boldsymbol{s}$.
\vfill
- *Ripley's*
\vfill
\vfill
- With CSR, $K(d)$
\vfill
- To estimate $K(d)$, 
\vfill
\vfill
- The empirical $K$ statistic is compared with $\pi d^2$. 
\vfill
- Similar to the $G$ and $F$ functions, edge correction is typically applied.
\newpage

### `spatstat` Intro
- The `spatstat` package is a comprehensive R package for point process data. It has a [website](http://spatstat.org) and a nice [vignette](https://cran.r-project.org/web/packages/spatstat/vignettes/getstart.pdf).
\vfill

- Consider a dataset with medieval grave site information.
```{r, echo = T}
data(grave)
summary(grave)
plot(grave)
```
\vfill
- Explore the `Fest()`, `Gest()`, and `Kest()` functions in `spatstat` and summarize the results for the `grave` dataset. (Also look at the `envelope` function for plots.)


\vfill

\newpage
- Next convert the point processes from early classes to `ppp` objects and explore `Fest()`/`Kest()`.

```{r, echo = T}
set.seed(04082019)
n <- rpois(4, 50)
x <- c(rbeta(n[1], 1, 1), rbeta(n[2], 1, 1), rbeta(n[3], 3, 1),rbeta(n[4], 3, 3))
y <- c(rbeta(n[1], 1, 1), rbeta(n[2], 1, 1), rbeta(n[3], 3, 1),rbeta(n[4], 3, 3))

comb.df <- data.frame(group = c(rep(1, n[1]), rep(2, n[2]), rep(3, n[3]), rep(4, n[4])), x = x, y = y)

df1 <- comb.df %>% filter(group ==1)
df2 <- comb.df %>% filter(group ==2)
df3 <- comb.df %>% filter(group ==3)
df4 <- comb.df %>% filter(group ==4)
```


### Estimating the intensity Function 

- With CSR, the intensity function is trivial 
\vfill
- __Discuss:__ given a realization of a point process, how could an intensity function be estimated?
\vfill
- *One option*
\vfill
\vfill
- *An alternative* 
\vfill
\vfill
- Now using the `plot(density(.))` function, plot and interpret the empirical intensity for the grave dataset along with the four synthetic examples.

\vfill